### 2014-11-01 BME 560/proposal.tex
New folder for BME 560; gitignored.

\documentclass{article}
\usepackage{tgbonum}
\usepackage{graphicx}
\title{Research Paper: Proposal}
\author{Rahul Krishna \\ 200020531}
\date{\today}
\begin{document}
	\maketitle
	\section{Introduction}
	\begin{list}{--}
	\item In the reseach paper for this term I plan on exploring Intensity Modulated Radiotherapy. 
	\item At the time the refered survey paper was complied, the technique was relatively new and implemented in only a few imaging centers around the world. In my survey, I'll include the results of some comparative studies and clinical trials that have been performed to show the efficacy of IMRT.
	\item The two types of tomotherapy-- linear and helical tomotheraphy. 
	\item Dose distribution
	\item Challenges in tomotherapy 
	\end{list}\end{document}