### 2014-12-08 BME 560/Paper.tex
Stuff added

\documentclass[10pt,journal,compsoc]{IEEEtran} %draftclsnofoot, 
\usepackage{graphicx}
\graphicspath{ {./_figures/} }
\setlength\fboxsep{1pt}
\setlength\fboxrule{1pt}
\usepackage{multicol}
\usepackage{bigstrut}
\usepackage{tabularx}
\usepackage{tabulary}
\usepackage{booktabs}
\usepackage{amsmath}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{hyperref}
\hypersetup{
  colorlinks = false,
  hidelinks = true
	}
\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage{times}
\usepackage{cite}

\begin{document}
  \markboth{BME 560 Medical Imaging: X-ray, CT, and Nuclear Methods. Fall, 
  2014}%
  {BME 560 Medical Imaging: X-ray, CT, and Nuclear Methods. Fall, 2014}

  \title{Intensity Modulated Radiation Therapy using Tomotherapy}
  \author{Rahul Krishna, 
  \IEEEauthorblockA{\normalsize {\textit{Dept. of Electrical and Computer 
  Engineering}\\
    North Carolina State University, Email: 
    \href{mailto:rkrish11@ncsu.edu}{{rkrish11@ncsu.edu}}}}}

	\IEEEcompsoctitleabstractindextext{%
  \begin{abstract}
    Tomotherapy is a form of 3D-conformal radiotherapy which involves the 
    delivery of intensity modulated radiation therapy (IMRT) using rotational 
    fan beam in a manner quite similar to that seen in modern CT scanners. The 
    modern tomotherapy systems have a couch and gantry which are in continuous 
    motion, closely resembeling the traditional helical-CT systems. This kind 
    of radiotherapy is hence called helical tomotherapy. Helical tomotherapy 
    delivers IMRT based on the images of the patient in the treatment position, 
    and these systems do this by acquiring CT images of the patient. This paper 
    is a discourse on the current state-of-the-art in the field, foucsing 
    primarily the technique's conceptual working and results of its clinical 
    implementation.

  \end{abstract}
  % IEEEtran.cls defaults to using nonbold math in the Abstract.
  % This preserves the distinction between vectors and scalars. However,
  % if the journal you are submitting to favors bold math in the abstract,
  % then you can use LaTeX's standard command \boldmath at the very start
  % of the abstract to achieve this. Many IEEE journals frown on math
  % in the abstract anyway. In particular, the Computer Society does
  % not want either math or citations to appear in the abstract.

  % Note that keywords are not normally used for peerreview papers.
  \begin{IEEEkeywords}
    Tomotherapy, Intensity Modulated Radiation Therapy.
  \end{IEEEkeywords}}
  \maketitle

  \section{Introduction}
	\IEEEPARstart{T}{he} past couple of decades have seen tremendous advances in 
	radiation oncology. The advent of faster and smaller computers has helped 
	transform the field of radiation therapy and treatment planning. The term 
	tomotherapy translates to ``slice therapy'', and it is derived from 
	tomography. Tomotherapy is a form of 3D Conformal Radiotherapy (3D-CRT), 
  which aims to conform the spatial distribution of the prescribed dose to a 
  target volume. The volume includes the cancerous cells plus a margin for 
  spatial uncertainties. While doing do it tries to minimize the dose to the
  surrounding normal structures. The tomotherapy system is designed to make use 
  of the tomographic 	reconstruction mathematics for treatment and also for 
  verification. The idea 	was conceived to tackle several major issues that 
  plagued radiation oncology. One of the primary issues was the limitation of 
  target dose that can be 	delivered due to the presence of neighboring 
  sensitive sturctures. One of the 	potential solutions to avoiding this was 
  to make use of multiple radiation 	beams with non-uniform beam intensities. 
  This method allows one to conform 	the radiation to the target region 
  thereby   sparing the sensitive normal 	structures which would otherwise 
  prevent   normal dose delivery to the affected 	region, see Fig. \ref{intro}.

  The setup has a 
  linear accelerator mounted on a rotating gantry as seen in CT 
  sacnners. The radiation is delivered to a patient in a helical way, obtained 
  by concurrent gantry rotation and couch/patient travel as shown in Fig. 
  \ref{fig1}. In general, the helical tomotherapy units have significantly 
  different radiation fields when compared to other radiotherapy systems. This 
  difference can be owed to the absence of absence of a flattening filter, a 
  thin target, an electron stopper, a beam hardener and a compact primary 
  collimator. This paper will briefly highlight the 
  radiation characteristics of the helical tomotherapy unit, especially in 
  parts where it differs from standard accelerators.

  The rest of this paper is 
  organized as follows. Section \ref{history} talks 
  about the history of tomotherapy. Section \ref{methods} has 3 subsections: 
  \ref{construction} talks about the construction of the tomotherapy unit, 
  \ref{radiation} discusses the radiation characteristics of helical 
  tomotherapy, and \ref{comparisions} brielfy examines the differences between 
  this techique and other conformal therapy techniques. Section \ref{clinical} 
  is a collection of some clinical applications where tomotherapy 
  has, and in some cases has not,  been effective. Finally, section 
  \ref{conclusions} presents concluding 
  remarks on the topic.

  \begin{figure}[htbp!]
    \includegraphics[width=\linewidth]{intro}
    \label{intro}
    \caption{An example of treatment planning in a modern 3D-CRT. In this 
    example, a 3D view of the patient, the PTV, spinal cord, and parotid 
    glands, and the 9 intensity modulated beams used to generate the IMRT dose 
    distribution can be seen. \textit{Image courtesy \cite{IMRT}}}
  \end{figure}

  \begin{figure*}[t!]
    \centering
  \includegraphics[width=0.89\linewidth]{fig1}
  \caption{A traditional tomotherapy unit, Ref \cite{Mackie1993}. (a) A block 
    diagram showing the ring gantry, and the radiation source}
  \label{fig1}

\end{figure*}

\begin{figure}[htbp!]
  \centering
  \includegraphics[width=0.9\linewidth]{fig2}
  \caption{``Running start-stop'' mechanism as proposed in Ref. 
    \cite{Mackie1993}. (a)--(c) }
  \label{fig2}

\end{figure}
	%----------------------------------
	% References
  %----------------------------------
  \section{History}
  \label{history}
  Tomotherapy was designed in the University of Wisconsin - Madison and 
  TomoTherapy Inc., Madison, WI. The first prototype investigated in this study 
  was installed at University of Wisconsin Hospital and Clinics in Madison, WI, 
  USA. The first paper on tomotherapy was submitted in 1992, Ref. 
  \cite{Mackie1993}. This paper described most of the details seen in the 
  modern sytems used today. This paper also introducted the idea of using a 
  continuously moving slip-ring gantry. The usage of fan-beam to form a 
  modulating beam, and a temorally modulating collimator assembly (this came 
  to be known as the `Binary Collimator') was also initially proposed in this 
  paper. It also recommended performing 
  intensity modulation without the use of a field flattening filter, thus 
  improving the spectrum across the beam.

  Several initial theories proposed in the paper could not be implemented. For 
  instance, to address the target cooling issue that is faced in modern photon 
  beam radiotherapy, the authors proposed engineering a stationary ``hoop'' 
  target with water supplied cooling; this could not be implemented due to the 
  increased collimation complexity. Also, in order to give a steep penumbra to 
  the target region, the authors of the original paper proposed a 
  ``start-stop'' mechanism where the collimator jaws defining the slice width 
  would be 
  close and later on as the patient traverses through the field the opposite 
  would happen. Fig. \ref{fig2}, taken from Ref. \cite{Mackie1993}, shows how 
  this was supposed to happen. But, this idea wasn't implementable either, 
  mainly because the technique required the jaws to be in movement during 
  treatment, optimization of beams near the ‘bow’ and the ‘stern’ of the 
  tumour may limit the amount of dose received.

  For a more comprehensive history the readers might want to refer to 
  \cite{Mackie2006}. The paper talks about the how this idea was 
  conceived and other technological advances led to its current state.

  \section{Methods and Materials}
  \label{methods}
  This section briefly discribes various aspects of helical tomography ranging 
  from it's construction and it's radiation characteristics to the differences 
  between it and other IMRT techniques.

  \subsection{Construction}
  \label{construction}
  \begin{figure}
    \centering
    \subfloat[][]{
      \includegraphics[width=\linewidth]{fig5}
      \label{fig4a}}\\
    \subfloat[][]{
      \includegraphics[width=\linewidth]{fig4}
      \label{fig4b}}
    \caption{\protect\subref{fig4a} Axial rotation of the fan beam as the 
      patient (represented 
      as a cylinder) traverses laterally (Along the y-axis). 
      \protect\subref{fig4b} A simplified view of the Helical 
      Tomotherapy 
      treatment system (at a gantry angle of 0$^\circ$) in use today. The 
      setup consists of a 6 MV linac to produce an x-ray beam which is 
      collimated down to a fan-beam using 2 moveable jaws (denoted as `y-jaws' 
      in the image). There is also a Binary Multileaf to modulate the beam 
      laterally. \textit{Images courtesy of \cite{Fenwick2004}}}
    \label{fig4}
  \end{figure}
  Figure \ref{fig4} show the basics of dose delivery in the helical tomotherapy 
  system. There is a short in-line 6 MV linear accelerator that rotates on a 
  ring gantry at a source-axis of 85 cms (instead of the usual 100 cms). The 
  patient couch is translated through the gantry bore in the y-direction as 
  shown in \ref{fig4}\protect\subref{fig4a}. The pitch is about 0.2 to 0.5 
  (Pitch is the distance traveled by the couch in one rotation, divided by the 
  field width in the direction of traversal). The pair of y-jaws in 
  \ref{fig4}\protect\subref{fig4b} are used to define the y-width of the beam. 
  These are usually preset to values like 1,2.5, or 5 cm before imaging. 
  Lateral modulation of the beam is achieved using the binary multileaf. This 
  system consists of 64 binary multileaves, each having a width of 6.25 mm 
  projecting onto an isocenter. These leaves transition rapidly between on 
  and off states, the duration of which is made use of to modualate the 
  intensity. Other parameters are tabulated in \ref{tab1}.
  % Table generated by Excel2LaTeX from sheet 'Sheet1'
  \begin{table}[t]
    \centering
    \caption{Helical Tomotherapy Characteristics}
    \begin{tabularx}{1\linewidth}{XX}
      \toprule[0.3mm]
      \textbf{Parameter} & \textbf{Value} \\
      \midrule
      Gantry Rotation angle & 
      7\textsuperscript{$\circ$} \\
      Projections per revolution & 51 \\
      Rotation velocity & 10 - 60 sec/rotation \\
      Time-per projection & $>$ 196 ms \\
      \bottomrule
    \end{tabularx}%
    \label{tab1}%
  \end{table}%                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             
  \begin{figure*}[htbp]
    \centering
    \subfloat[][]{
      \includegraphics[width=0.5 \linewidth]{fig6}
      \label{radchara}}
    \subfloat[][]{
      \includegraphics[width=0.5 \linewidth]{fig7}
      \label{radcharb}}
    \caption{\protect\subref{radchara} Shows the photon fluence profile of the 
    tomotherapy treatment beam. Notice the characteristic conical shape of the 
    tomotherapy profile (solid red line) and the flat shape of the conventional 
    linac IMRT system (dashed blue line). 
    \protect\subref{radcharb} Average photon beam spectra of the helical 
    tomography system in two operational modes: imaging (dashed green) and 
    treatment mode (solid red). \textit{Image courtesy of \cite{Jeraj2004}}}
    \label{fig5}
  \end{figure*}
  \subsubsection{A note on Serial Tomotherapy}
  Serial tomography is another variant of tomotherapy that has been in use for 
  sometime now. The first clinical implementation occured as early as 1994. The 
  setup consists of a binary multileaf collimator attached to a conventional 
  linear accelerator. The collimator comprises of two opposing banks of 20 
  tungsten leaves which are driven pneumatically to lie either within or just 
  outside the fan beam radiation. The fan beam itself is modified by arranging 
  the leaves to lie within the radiation field for a variable amount of time.

  This can also be achieved by slightly modifying the conventional linear 
  accelerator by adding a binary multileaf collimator to the head of the linac. 
  Axial tomotherapy dose distributions are delivered slice-by-slice, with 
  patients being sequentially and discretely translated through the linac 
  gantry rotational plane between slices, helical distributions are delivered 
  without interruptions; patients are translated smoothly through the bore of 
  the machine as its gantry continuously and synchronously rotates, the 
  therapy equivalent of spiral computed tomography \cite{Fenwick2006}. 
  \subsection{Radiation Characteristics}
  \label{radiation}
  This section aims to briefly highlight the key differences in radiation 
  characteristics of helical tomography and other forms of radiotherapy. One of 
  the key differences is the lack of flattening filter, which makes the dose 
  more uniform are greater depths. As a result of this, the photon fluence 
  profile is shaped differently when compared to a traditional radiotherapy 
  system, as can be seen in Fig. \ref{radchara}. The conical shape of the 
  profile implies that there will be an increased average dose rate and thus 
  reduces the imaging time.

  Figure \ref{radcharb} shows the differences in the photon spectrum between 
  the 
  treatment and imaging modes of helical tomotherapy. Notice that the in the 
  imaging mode the spectrum has a lower energy when compared to treatment mode; 
  this is done so as to obatin better images by improving the contrast. A more 
  comprehensive comparision is documented in Ref. \cite{Jeraj2004}

  \subsection{Comparisions between IMRT techniques}
  \label{comparisions}
    \begin{table*}[t!]
      \centering
      \caption{A comparision between standard and non-standard IMRT delivery 
      systems (Ref. \cite{Fenwick2006})}
      \small{
      \begin{tabulary}{\linewidth}{LLLLLLL}
        \toprule[0.04cm]
        \textbf{System} & \textbf{Gantry} & \textbf{ Beam Geometry} & 
        \textbf{Collimated} & \textbf{Modulated} & \textbf{Imaging} & 
        \textbf{Gating
          Potential} \\\toprule[0.04cm]

        Conventional linac plus multileaf & C-Arm & Non-coplanar cone beam & 
        Jaws 
        with conventional
        multileaf & Jaws with conventional
        multileaf & Portal image, fluoroscopy, kvCT, 
        infrared 
        reflectors &
        Breath-hold, 
        beam trigger, multileaf tracking \bigstrut\\\midrule
        Conventional linac minus multileaf & C-Arm & Non-coplanar cone beam & 
        Jaws with/without tertiary attenuator
        & Jaws with/without tertiary attenuator
        & Portal image, fluoroscopy, kvCT, infrared 
        reflectors & Breath-hold,
        beam trigger,
        jaw tracking \bigstrut\\\midrule
        Serial tomotherapy & C-Arm & Fanbeam, coplanar, sequential,
        indexed trajectory & Jaws with binary
        multileaf & Binary multileaf & Portal image, 
        fluoroscopy, kvCT, infrared reflectors & Breath-hold
        during each
        slice \bigstrut\\\midrule
        Helical tomotherapy & Ring  &  Fanbeam,
        coplanar, helical
        trajectory & Jaws with binary
        multileaf & Binary multileaf & MVCT, infrared 
        reflectors & Delivery
        synchronized
        with breathing
        cycle \bigstrut \\\midrule
        Robotic linac & Robotic Arm &  Pencil-beam,
        noncoplanar & Circular collimator & Superposition of pencilbeams by 
        robotic arm & Biplanar radiography, infrared 
        reflectors & Beam 
        trigger,
        robotic
        tracking \bigstrut \\
        \bottomrule
      \end{tabulary}}%

      \label{tab:compare}%
    \end{table*}%
  Tomotherapy is different from other standard IMRT techiniques such as 
  conformal radiotherapy (CRT) techniques and non-standard IMRT techniques like 
  Cyber knife.   Standard CRT techniques generate highly conformal radiotherapy 
  using radiation fields generated in a few fixed-angle positions. These are 
  generated by linear accelerators (aka. LINAC) which are then modulated by 
  conventional leaf collimators operated in either dynamic mode or the standard 
  step-and-shoot mode. CyberKnife is a modern radiotherapy system developed by 
  Accuray, Inc., primarily for radiosurgery. It makes use of a highly precise 
  robotic arm to delivery the doses. The arm is computer driven to achieve very 
  high precision, it is mounted with a compact 6-MV LINAC whose orientation is 
  determined by 6 different mechanical subsystems. This robotic arm can be 
  moved with a much greater generality than most conventional sytsems.

  Table \ref{tab:compare} highlights the key differences in various systems. 
  For the 
  sake the completeness, the authors of Ref. \cite{Fenwick2006} have 
  dicotomized the conventional system into 2 types, with and without a 
  multileaf collimator, and the tomotherapy into serial and parallel 
  tomotherapy.
  \section{Clinical Applications}
  \label{clinical}
  This section highlights few of the many clinical applications of tomotherapy. 
  \subsection{Total Body Irradiation (TBI)}
  Tomotherapy can be used in establishing total body irradiation (TBI) to gain 
  greater control over the dose distribution and to spare organs that may be at 
  risk.

  One of the disadvantages of using Helical Tomography (HT) for TBI is that it 
  can't be used when the body length of the patient exceeds 145 cms. For 
  patients who are taller than this, a better solution needs to be deviced.
  \subsection{Whole brain helical Tomotherapy}
  \subsection{Inoperable Lung Cancer}
  \subsection{Rectal Cancer}
  \section{Conclusion}
  \label{conclusions}
  \bibliography{Paper}{}
  \bibliographystyle{IEEEtran}

\end{document}